---
title: "EJ tool: All indicators"
author: "Amanda Gorton, MPCA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Overview of data indicators + initial set up 

This script will contain the cleaning and percentile ranking for all of the data indicators used in the EJ tool.  

I also have these split by individual indicators in the shared EJ tool folder, found here:  
"X:\Agency_Files\Data_Services\DAU\Analyses\EJ\EJ Score - Metrics for Scoring\EJ Prioritization Tool Version 1\Data Files - Organized"  

## Load packages

```{r load packages}
library(readxl)
library(tidyr)
library(janitor)
library(dplyr)
library(tidyverse)
```

## Import data

Where possible, I have indicated the source of each indicator. Please refer to the methodology and written report for the EJ tool for more information  

```{r load data}
# Import CDC data for entire US only once - afterwards use subset for MN below
#PLACES <- read.csv("PLACESData_CDC.csv")

```

# Data clean-up and manipulation

## CDC PLACES

### Filtering, shortening, and cleaning up data

```{r subset CDC PLACES data}
# Subset dataset so that it's only for MN
MN <- PLACES %>% 
  filter(StateAbbr == 'MN')

# Drop unnecessary columns
MN_abbr <- MN %>% 
  select(-c(LocationName, DataSource, Category, Data_Value_Footnote_Symbol, Data_Value_Footnote, CategoryID, MeasureId, DataValueTypeID, Short_Question_Text))

# Export this reduced dataset for future use
write.csv(MN_abbr, "CDCPlaces_MN.csv")

# Look at the health measures to figure out which one are for asthma and heart disease
# "Current asthma among adults aged >=18 years"   
# "Coronary heart disease among adults aged >=18 years"  
unique(MN$Measure)

asthma <- MN_abbr %>% 
  filter(Measure == 'Current asthma among adults aged >=18 years')

heartdisease <- MN_abbr %>% 
  filter(Measure == "Coronary heart disease among adults aged >=18 years")

# Export these to send to MDH for comparison
#write.csv(asthma, "CDCPlaces_asthma.csv")
#write.csv(heartdisease, "CDCPlaces_heartdisease.csv")

```

### Calculate percentile ranking

```{r}

# For asthma and heart disease

asthma_percentiles <- asthma %>%
  rename(geoid = LocationID) %>%
  mutate(percentile_rank = percent_rank(Data_Value))

heartdisease_percentiles <- heartdisease %>%
  rename(geoid = LocationID) %>%
  mutate(percentile_rank = percent_rank(Data_Value))


# Ranked percentiles for percentage of adults with asthma, by tract
write.csv(asthma_percentiles, "asthma_percentiles.csv")

# Ranked percentile for percentage of adults with heart disease, by tract
write.csv(heartdisease_percentiles, "heartdisease_percentiles.csv")

```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
