---
title: "EJ Scoring Tool - ACS Data"
author: "Amanda Gorton, MPCA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Set up

## Load packages

```{r}
library(tidycensus)
library(keyring)
library(tidyverse)
```

## Load API key

Might need to do this every time - need to check

```{r, eval = FALSE}
census_api_key(key_get("census api"))
```

## ACS variables

Use the code below to browse the variable names in the subject tables

```{r, eval = FALSE}

# Use 'filter' to change what I am specifically looking for, either by including a particular number for the variable (name), or searching for words  in the label (e.g. Unemploy, Education, etc)

# Income - Federal Poverty Level
acs.income <- load_variables(year = 2020, dataset = "acs5/subject") %>% 
  filter(str_detect(name, "S1701"))

# Educational attainment
acs.educ.percent <- load_variables(year = 2020, dataset = "acs5/subject") %>% 
  filter(str_detect(name, "S1501_C02"))

# Age + sex
acs.subject <- load_variables(year = 2020, dataset = "acs5/subject") %>% 
  filter(str_detect(name, "S0101_C01"))

# Limited English
acs.english <- load_variables(year = 2020, dataset = "acs5/subject") %>% 
  filter(str_detect(name, "S1602"))

# Race
acs.race <- load_variables(year = 2020, dataset = "acs5") %>% 
  filter(str_detect(name, "B03002"))

#B03002_001 - Total population
#B03002_004 - Black or African American alone
#B03002_005 - American Indian and Alaska Native alone
#B03002_007 - Native Hawaiian and Other Pacific Islander
#B03002_008 - Some other race
#B03002_009 - Two or more races
#B03002_012 - Hispanic or Latino


```


### Type of tables 

#### Subject tables (S)

Pretabulated estimates and percentages. Begin with the letter S.  
https://www.census.gov/acs/www/data/data-tables-and-tools/subject-tables/ 

#### Detailed Tables (B or C)  

Detailed Tables are the most comprehensive ACS tables  
"B" = base tables  
"C" = collapsed tables, which cover the same topics but with fewer categories  
https://data.census.gov/cedsci/all?d=ACS%205-Year%20Estimates%20Detailed%20Tables  

### Socioeconomic categories for EJ scoring tool  

These variables come from pre-existing categorizations that are used for the EJ Scoring Tool.  
Some of them are multiple variables summed together, while others are stand-alone variables. 

**Age** 
Number of people 5 and under AND 65+    

Table: S0101 - Age and Sex, Field names = C01_002, C01_030

**Race**
Number of people that are non-white.    
Add up all race categories that are non-white    

Table: B03002 - Hispanic or Latino Origin by Race  
https://censusreporter.org/tables/B03002/   

**Educational attainment**  
Percentage of the population _without_ a high school diploma  
Use the variable of people with a high school diploma or higher to calculate this.   

Table: S1501 - Educational Attainment  
Field name = C02_14 (Percent) or C01_014 (Count); Population > 25 years, high school or higher  
Total pop field name = C01_006

**Limited english speaking**
Percentage of the population with limited english-speaking households  

Table: S1602 - Limited English Speaking Households  
Field name: C04_001 (Percent) or C03_001 (Count); Limited English speaking households, All households  
All households field name: C01_001

**Income**
Number of people with income at or below 200% of the federal poverty level  

Table: S1701 - Poverty Status in the Past 12 Months
Field name: C02_042 (Count) 
Total pop field name = C01_001


# Import ACS variables + manipulate


```{r, eval = FALSE, echo = FALSE, results = HIDE}

# Mix of raw counts and/or percentages.  
#The code below brings in the variables that Quinn has used previously


acs_ejscoring <- get_acs(
  state = "MN",
  geography = "tract",
  variables = c(Age_TotalPop = "S0101_C01_001", 
                Age_Under5 = "S0101_C01_002",
                Age_Over65 = "S0101_C01_030",
                Percent_Educ_Attain_Over25_HighSchool = "S1501_C02_014",
                Percent_LimitedEnglish = "S1602_C04_001",
                Income_TotalPop_FedPovLevel = "S1701_C01_001", 
                Income_Under200_FedPovLevel =  "S1701_C02_042", 
                Race_TotalPop = "B03002_001",
                Race_BlackAfricanAmerican = "B03002_004",
                Race_AmericanIndianAlaskaNative = "B03002_005",
                Race_Other = "B03002_008",
                Race_TwoOrMore = "B03002_009",
                Race_HispanicLatino = "B03002_012"
                ), 
  geometry = TRUE, #Adding in the spatial component
  year = 2020
)


#B03002_001 - Total population
#B03002_004 - Black or African American alone
#B03002_005 - American Indian and Alaska Native alone
#B03002_007 - Native Hawaiian and Other Pacific Islander
#B03002_008 - Some other race
#B03002_009 - Two or more races
#B03002_012 - Hispanic or Latino
```

```{r, eval = FALSE, echo = FALSE}

#Taking a quick glance, it seems as though 'TotalPop' may be the same across multiple variables - this makes sense, as they are all at the same level (i.e. tract). Let's check this using filter + dplyr.  

acs.totalpop <- acs_ejscoring %>% 
  filter(str_detect(variable, "TotalPop"))
```

## Import all ACS variables as counts

Import ACS variables, using estimates/counts for all and adding in total population denominator to normalize counts 

```{r, results = HIDE}
acs_counts <- get_acs(
  state = "MN",
  geography = "tract",
  variables = c(Age_Under5 = "S0101_C01_002",
                Age_Over65 = "S0101_C01_030",
                EducAttain_Over25_HighSchool = "S1501_C01_014",
                LimitedEnglish = "S1602_C03_001",
                Income_Under200_FedPovLevel =  "S1701_C01_042", 
                Race_BlackAfricanAmerican = "B03002_004",
                Race_AmericanIndianAlaskaNative = "B03002_005",
                Race_Other = "B03002_008",
                Race_TwoOrMore = "B03002_009",
                Race_HispanicLatino = "B03002_012"
                ), 
  summary_var = c(TotalPop = "S0101_C01_001"),
  #geometry = TRUE, #Adding in the spatial component
  year = 2020
)


```

## Normalize and convert counts to percentages

```{r}
acs_percent <- acs_counts %>%
  mutate(percent = 100*(estimate/summary_est)) %>%
  select(GEOID, NAME, variable, estimate, moe, percent) # Select only certain columns to retain
```


## Calculate socioeconomic indcators from ACS variables 

Age = Age_Under5 + Age_Over65  
Race = Race_BlackAfricanAmerican, Race_AmericanIndianAlaskaNative, Race_Other, Race_TwoOrMore, Race_HispanicLatino
Educational Attainment = (EducaAttain_Over25_HighSchool - 1)
Limited english speaking = LimitedEnglish
Income = Income_Under200_FedPovLevel


```{r}

# First, reassign all the estimates for each variable into the different indicator groups

# https://walker-data.com/census-r/wrangling-census-data-with-tidyverse-tools.html
# See section 3.3.2 Tabulating new groups

ej_ind <- acs_percent %>% 
    group_by(NAME) %>%
    mutate(indicator = case_when(
      variable == c("Race_BlackAfricanAmerican", "Race_AmericanIndianAlaskaNative", "Race_Other", "Race_TwoOrMore", "Race_HispanicLatino") ~ "race",
      variable == "LimitedEnglish" ~ "limited_english",
      variable == "Income_Under200_FedPovLevel" ~ "income",
      variable == "EducAttain_Over25_HighSchool" ~ "education",
      TRUE ~ "age"
    ))

```

```{r}

# Next, summarize percentages into the new indicator groupings 

ej_sums <- ej_ind %>%
  group_by(GEOID, NAME, indicator) %>%
  summarize(percent = sum(percent))

```

```{r, eval = FALSE, echo = FALSE}
# Have to ignore the geometry to do the summary 

ej_ind_sum <- ej_indicators %>%
  as.data.frame() %>% # Include this to remove sticky geometry
  group_by(GEOID, indicator) %>%
  summarize(percent = sum(percent))

#st_drop_geometry()

# Add back in geometry

# Doesn't work
ej_ind_sum_sf <-
  st_join(ej_ind_sum, ej_ind %>%
  select(NAME), by = GEOID)

# Doesn't work
ej_ind_sum_sf <-
  merge(ej_ind_sum, ej_ind, by = "GEOID")


```

## Assign percentiles for each indicator

Need to assign percentiles for each indicator, so that each tract has a percentile associated with each of the five indicators

```{r}
ej_percentiles <- ej_sums %>%
  group_by(indicator) %>%
  mutate(percent_rank = rank(percent)/length(percent))
```

## Pivot data wider

In order to join these data with the other data types used in the EJ Score Tool, need to pivot data wider

```{r}

ej_percentiles_wide <- ej_percentiles %>%
  select("GEOID", "percent_rank", "indicator") %>%
  pivot_wider(names_from = "indicator", values_from = c("percent_rank"))

```

# Export data

Ranked percentiles for each indicator, with one row per tract

```{r}
write.csv(ej_percentiles_wide, "ACS_20162020_ejindicators_percentiles.csv")
```

