---
title: "EJ tool: All indicators"
author: "Amanda Gorton, MPCA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Combining ranked indicators 

Goal: Combine all the ranked indicators into a single file to be used in ArcGIS for creating the EJ tool

**What do I need to do to accomplish this?**
1) Import all the ranked data (i.e. percentiles that were already ranked)
2) Use janitor for consistency in case and names
3) Drop unnecessary columns, and rename points and score columns so that they are associated with the indicator type
4) Make sure percentile is a percent (i.e out of 100) rather than a proportion; drop to 2 deimcals
5) Merge based on GEOID
6) Convert to a .shp file for ArcGIS

## Import ranked percentile data and cleanup

```{r}
# Make sure pointing to the 'output' directory - will probably have to move this to the XDrive at some point
# H:\R\ej-tool\output

cleanup <- read.csv('H:/R/ej-tool/output/cleanup_sites_percentiles.csv') #Why does this work!? GAH!
sw <- read.csv('H:/R/ej-tool/output/solid_waste_percentiles.csv')
groundwater <- read.csv('H:/R/ej-tool/output/groundwater_threats_percentiles.csv')
hw <- read.csv('H:/R/ej-tool/output/hazardouswaste_percentiles.csv')
impaired <- read.csv('H:/R/ej-tool/output/impairedwaters_percentiles.csv')

```

Let's take a look to see what are all our column names and styles. This will tell me which ones I need to keep/drop or change names  
Also check which percentiles need to be multiplied by 100
```{r, eval = FALSE}

names(cleanup) # Keep geoid, cleanup_points, and percentile_rank. Add cleanup to percentile name
names(sw) # Add 'sw' in front of percentile_rank. Keep geoid, sw_points, and percentile_rank. 
names(groundwater) # Rename groundwater_score to groundwater_points. Keep geoid and percentile_rank. Add groundwater to percentile name
names(hw) # Use janitor to change to lower case. Keep hw_points and percentile_rank. Add hw to percentile rank name
names(impaired) # Keep geoid, impaired_score, percentile_rank. Add impaired to percentile_rank name and change score to points

head(cleanup) # x100
head(sw) #all good
head(groundwater) # x100
head(hw) #all good
head(impaired) #all good
```

Drop and rename columns as necessary
```{r dropping and renaming columns}

# Cleanup sites
cleanup <- cleanup %>%
  mutate(percentile_rank = percentile_rank*100) %>%
  rename(cleanup_percentile = percentile_rank) %>%
  mutate(across(c(cleanup_percentile), round, 2)) %>%
  select("geoid", "cleanup_points", "cleanup_percentile")

# Solid waste
sw <- sw %>%
  rename(sw_percentile = percentile_rank) %>%
  mutate(across(c(sw_percentile), round, 2)) %>%
  select("geoid", "sw_points", "sw_percentile")

# Groundwater
groundwater <- groundwater %>%
  mutate(percentile_rank = percentile_rank*100) %>%
  rename(groundwater_percentile = percentile_rank, groundwater_points = groundwater_score) %>%
  mutate(across(c(groundwater_percentile), round, 2)) %>%
  select("geoid", "groundwater_points", "groundwater_percentile")

# Hazardous waste
hw <- hw %>%
  clean_names() %>%
  rename(hw_percentile = percentile_rank) %>%
  mutate(across(c(hw_percentile), round, 2)) %>%
  select("geoid", "hw_points", "hw_percentile")

# Impaired waters
impaired <- impaired %>%
  rename(impaired_percentile = percentile_rank, impaired_points = impairment_score) %>%
  mutate(across(c(impaired_percentile), round, 2)) %>%
  select("geoid", "impaired_points", "impaired_percentile")



```

Merge the indicators based on geoid
```{r merge on geoid}

# Add all the dataframes to a list
indicators <- list(cleanup, groundwater, hw, impaired, sw)

# Merge by geoid
all <- indicators %>%
  reduce(full_join, by = 'geoid')

```

Export for now
```{r}
write.csv(all, "allindicators_percentiles.csv")

```


# Overview of data indicators + initial set up 

This script will contain the cleaning and percentile ranking for all of the data indicators used in the EJ tool.  

I also have these split by individual indicators in the shared EJ tool folder, found here:  
"X:\Agency_Files\Data_Services\DAU\Analyses\EJ\EJ Score - Metrics for Scoring\EJ Prioritization Tool Version 1\Data Files - Organized"  

## Load packages

```{r load packages}
library(readxl)
library(tidyr)
library(janitor)
library(dplyr)
library(tidyverse)
```

## Import data

Where possible, I have indicated the source of each indicator. Please refer to the methodology and written report for the EJ tool for more information  

```{r load data}
# Import CDC data for entire US only once - afterwards use subset for MN below
#PLACES <- read.csv("PLACESData_CDC.csv")

```

# Data clean-up and manipulation: calculatuing percentiles for each indicator

## CDC PLACES

### Filtering, shortening, and cleaning up data

```{r subset CDC PLACES data}
# Subset dataset so that it's only for MN
MN <- PLACES %>% 
  filter(StateAbbr == 'MN')

# Drop unnecessary columns
MN_abbr <- MN %>% 
  select(-c(LocationName, DataSource, Category, Data_Value_Footnote_Symbol, Data_Value_Footnote, CategoryID, MeasureId, DataValueTypeID, Short_Question_Text))

# Export this reduced dataset for future use
write.csv(MN_abbr, "CDCPlaces_MN.csv")

# Look at the health measures to figure out which one are for asthma and heart disease
# "Current asthma among adults aged >=18 years"   
# "Coronary heart disease among adults aged >=18 years"  
unique(MN$Measure)

asthma <- MN_abbr %>% 
  filter(Measure == 'Current asthma among adults aged >=18 years')

heartdisease <- MN_abbr %>% 
  filter(Measure == "Coronary heart disease among adults aged >=18 years")

# Export these to send to MDH for comparison
#write.csv(asthma, "CDCPlaces_asthma.csv")
#write.csv(heartdisease, "CDCPlaces_heartdisease.csv")

```

### Calculate percentile ranking

```{r}

# For asthma and heart disease

asthma_percentiles <- asthma %>%
  rename(geoid = LocationID) %>%
  mutate(percentile_rank = percent_rank(Data_Value))

heartdisease_percentiles <- heartdisease %>%
  rename(geoid = LocationID) %>%
  mutate(percentile_rank = percent_rank(Data_Value))


# Ranked percentiles for percentage of adults with asthma, by tract
write.csv(asthma_percentiles, "asthma_percentiles.csv")

# Ranked percentile for percentage of adults with heart disease, by tract
write.csv(heartdisease_percentiles, "heartdisease_percentiles.csv")

```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
