---
title: 'EJ tool: Solid waste data'
author: "Amanda Gorton, MPCA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Set-up and overview

### Data description

**Combine the following data into a single table**
- compost - omit for now until Quinn fixes these data
- gas capture
- landfill
- leachate
- tires
- transfer areas
- waste to energy

**Fields in common**
- AI name
- AI ID
- permit number
- latitude
- longitude
- waste type
- waste amount
- waste units

## Load packages and import data 
```{r import packages}
library(tidyr)
library(janitor)
library(dplyr)
library(tidyverse)
library(readxl)
```

```{r import data}
compost <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Compost Facilities with Coordinates and Points.xlsx", sheet = "CompostSitesPointsCoordinates")

gas <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Gas Capture Facilities.xlsx", sheet = "Gas Capture Coordinates Points")

landfill <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Landfill or Construction, Demolition and Inert Debris Waste Disposal Sites.xlsx", sheet = "Percentile Ranking Histogram")

leachate <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Leachate Facilities.xlsx", sheet = "Leachate Coordinates Points")

tires <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Tire Waste Facilities.xlsx", sheet = "Tire Waste with Converted Units")

transfer <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Transfer Area Facilities.xlsx", sheet = "Full Data with Converted Units")

wte <- read_excel("X:\\Agency_Files\\Data_Services\\DAU\\Analyses\\EJ\\EJ Score - Metrics for Scoring\\EJ Prioritization Tool Version 1\\Data Files - Organized\\Solid waste\\Solid Waste Waste-to-Energy Facilities.xlsx", sheet = "WTE Facilities 2021")

```

## Clean-up and formating datasets

**What needs to be done to each dataframe?**

1. Clean up column names to remove spaces, upper case
2. Rename columns names to common names across datasets
3. Make new column with waste type
4. Drop extraneous columns

### Compost

**Keep**: Latitude (rename), Longitude (rename), Annual capacity (rename), Master ai name (rename), Permit number, Units (rename), Master ai id (rename)
**Add**: waste_type = compost

```{r compost}

compost <- compost %>%
  clean_names %>%
  rename(permit_number = permit, waste_amount = annual_capacity, ai_name = master_ai_name, latitude = latitude_solid_waste_permit_si_avg, longitude = longitude_solid_waste_permit_si_avg, units = units_desc, ai_id = master_ai_id) %>%
  mutate(waste_type = 'compost') %>%
  mutate_at(c('latitude', 'longitude', 'ai_id'), as.numeric) %>%
  select("latitude", "longitude", 'permit_number', 'ai_name', 'ai_id', 'units', 'waste_amount', 'waste_type')


```


### Gas capture

**Keep**: Latitude, Longitude, Units, Total Gas Collected (rename), Permit Number (rename)
**Add**: waste_type = gas capture

```{r gas capture}

gas <- gas %>%
  clean_names %>%
  rename(permit_number = permit_number_gas_capture_for_tableau, waste_amount = total_gas_collected_cubic_meters) %>%
  mutate(waste_type = 'gas capture') %>%
  mutate_at(c('latitude', 'longitude'), as.numeric) %>%
  select("latitude", "longitude", "units", 'permit_number', 'waste_amount', 'waste_type')

```

### Landfill

**Keep**: Latitude, Longitude, Permit number, Facility (rename), Overall annual cubic yards of waste (rename)
**Add**: waste_type = landfill

```{r landfill}

landfill<- landfill %>%
  clean_names %>%
  rename(ai_name = facility, waste_amount = overall_annual_cubic_yards_of_waste) %>%
  mutate(waste_type = 'landfill or construction, demo and inert', units = 'cubic yards') %>%
  mutate_at(c('latitude', 'longitude'), as.numeric) %>%
  select('ai_name', "latitude", "longitude", 'waste_amount', 'waste_type', 'permit_number', 'units')

```



### Leachate

**Keep**: Latitude, Longitude, Units, Master AI ID (rename), Total Annual Leachate (rename)
**Add**: waste_type = leachate

```{r}
leachate <- leachate %>%
  clean_names %>%
  rename(ai_id = master_ai_id, waste_amount = total_annual_leachate, ) %>%
  mutate(waste_type = 'leachate') %>%
  mutate_at(c('latitude', 'longitude', 'ai_id', 'waste_amount'), as.numeric) %>%
  select('ai_id', "latitude", "longitude", "units", 'waste_amount', 'waste_type')


```

### Tires

**Keep**: Final Tons (rename), Permit Number, Master AI ID (rename), Facility Name, Waste action description (rename), Units
**Add**: waste_type = tires

```{r tires}
tires <- tires %>%
  clean_names %>%
  rename(ai_id = master_ai_id, ai_name = facility_name, waste_amount = final_tons, waste_action = waste_action_description) %>%
  mutate(waste_type = 'tires') %>%
  mutate_at('ai_id', as.numeric) %>%
  select('ai_id', 'units', 'waste_amount', 'waste_type', 'permit_number', 'ai_name', 'units')

```

### Transfer areas

**Keep**: Amount tons final (rename), Permit number, Master AI ID (rename), Facility name, Waste action description (rename), Units
**Add**: waste_type = transfer areas

```{r transfer areas}

transfer <- transfer %>%
  clean_names %>%
  rename(ai_id = master_ai_id, waste_amount = amount_tons_final, waste_action = waste_action_description, ai_name = facility_name) %>%
  mutate(waste_type = 'transfer areas') %>%
  mutate_at('ai_id', as.numeric) %>%
  select('ai_id', 'units', 'waste_amount', 'waste_type', 'permit_number', 'ai_name')


```

### Waste to energy 

**Keep**: Amount (rename), Units, Permit number, Master AI ID (rename), Facility name, Waste action description (rename)
**Add**: waste_type = waste to energy

```{r waste to energy}
wte <- wte %>%
  clean_names %>%
  rename(ai_id = master_ai_id, waste_amount = amount, waste_action = waste_action_description, ai_name = facility_name) %>%
  mutate(waste_type = 'waste to energy') %>%
  mutate_at('ai_id', as.numeric) %>%
  select('ai_id', 'units', 'waste_amount', 'waste_type', 'permit_number', 'ai_name')

```

## Merging datasets together

Next, I want to combine compost, gas capture, landfill, leachate, tires, transfers areas and waste to energy into a single dataframe  

Need to replace NAs with blanks, but not sure how to get it to work right now

```{r bind rows}
# Combine them together
smooshy <- bind_rows(compost, gas, landfill, leachate, tires, transfer, wte)

# This doesn't quite work
test <- replace(smooshy, is.na(smooshy), "")

```

## Export data

Combined SW dataset
```{r export sw data}

write.csv(smooshy, "solidwaste_combined.csv")

```

